### 2024-09-21 paper/paper.qmd
cleaned imported data

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-09-21 scripts/00-simulate_data.R

@@ -30,6 +30,5 @@ data <-
     number_of_marriage = rpois(n = number_of_dates, lambda = 10)
   )
 
-
 #### Write_csv
 write_csv(data, file = "data/raw_data/simulated.csv")

### 2024-09-21 scripts/02-data_cleaning.R

@@ -1,44 +1,29 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Cleans the city council attendance data 
+# Author: Wendy Yuan
+# Date: 27 September 2024
+# Contact: w.yuan@mail.utoronto.ca
 
 #### Workspace setup ####
 library(tidyverse)
+library(dplyr)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("data/raw_data/unedited_data.csv")
 
 cleaned_data <-
   raw_data |>
   janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+  filter(committee == "City Council") |> 
+  #change from Y = present to 1 and N = absent to 0
+  mutate(present = ifelse(present == "Y", 1, 0)) |>
+  #removing day from date 
+  separate(col = session_date,
+           into = c("session_year", "session_month"),
+           sep = "-") |> 
+  #selects columns of interest
+  select(first_name, last_name, committee, 
[…]

### 2024-09-21 scripts/03-test_data.R

@@ -1,15 +1,17 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Tests the city council attendance data
+# Author: Wendy Yuan
+# Date: 27 September 2024
+# Contact: w.yuan@mail.utoronto.ca
 
+#### Tests raw data ####
+raw_data <- read_csv("data/raw_data/unedited_data.csv")
 
-#### Workspace setup ####
-library(tidyverse)
-# [...UPDATE THIS...]
+# Test for NAs
+all(is.na(raw_data))
 
-#### Test data ####
+#### Tests cleaned data ####
+cleaned_data <- read_csv("data/analysis_data/cleaned_data.csv")
+
+# Test for NAs
+cleaned_data$committee == "City Council"

### 2024-09-21 scripts/00-simulate_data.R
simulated data

@@ -10,25 +10,19 @@ library(tidyverse)
 #### Simulate data ####
 set.seed(304)
 
-# Define the start and end date
-start_date <- as.Date("2018-01-01")
-end_date <- as.Date("2023-12-31")
+observations <- 100
+year <- c(2022:2024)
+months <- c(1:12)
+type <- c("morning", "afternoon")
 
-# Set the number of random dates you want to generate
-number_of_dates <- 100
+session_type <- sample(type, observations, replace = TRUE)
+session_year <- sample(year, observations, replace = TRUE)
+session_month <- sample(months, observations, replace = TRUE)
+attendance <- rbinom(observations, 1, 0.75)
 
-data <-
-  tibble(
-    dates = as.Date(
-      runif(
-        n = number_of_dates,
-        min = as.numeric(start_date),
-        max = as.numeric(end_date)
-      ),
-      origin = "1970-01-01"
-    ),
-    number_of_marriage = rpois(n = number_of_dates, lambda = 10)
-  )
+sim_data <- 
+  data.frame(session_type, session_year, session_month, attendance)
 
 #### Write_csv
-write_csv(data, file = "data/raw_data/simulated.csv")
+write_csv(sim_data, file = "data/simulated_data/simulated_data.csv")
+

### 2024-09-21 scripts/02-data_cleaning.R

@@ -26,4 +26,4 @@ cleaned_data <-
          session_year, session_month, session_type, present)
 
 #### Save data ####
-write_csv(cleaned_data, "data/analysis_data/cleaned_data.csv")
+write_csv(cleaned_data, "data/cleaned_data/cleaned_data.csv")

### 2024-09-21 scripts/03-test_data.R

@@ -11,7 +11,7 @@ raw_data <- read_csv("data/raw_data/unedited_data.csv")
 all(is.na(raw_data))
 
 #### Tests cleaned data ####
-cleaned_data <- read_csv("data/analysis_data/cleaned_data.csv")
+cleaned_data <- read_csv("data/cleaned_data/cleaned_data.csv")
 
 # Test for NAs
 cleaned_data$committee == "City Council"

### 2024-09-21 paper/paper.qmd
produced some graphs in quarto pdf

@@
 author: 
-  - First author
-  - Another author
+  - Wendy Yuan
 thanks: "Code and data are available at: LINK."
-date: today
+date: "27 September 2024"
 date-format: long
@@
 library(tidyverse)
-library(palmerpenguins)
+library(dplyr)
+
+cleaned_data <- read_csv(here::here("data/cleaned_data/cleaned_data.csv"))
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
-
-The remainder of this paper is structured as follows. @sec-data....
-
 # Data {#sec-data}
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-session_type
+#| fig-cap: Comparison of Attendance in Morning, Afternoon, vs Evening Session
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-```{=tex}
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-```
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+ggplot(cleaned_data, aes(x = session_type, fill = present)) +
+  geom_bar(position = "dodge2") 
 ```
@@
 ```{r}
+#| label: fig-session_month
+#| fig-cap: Comparison of Attendance over Different Months
 #| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+cleaned_data |> 
+  ggplot(aes(x = session_month, fill = present)) +
+  geom_bar(position = "dodge2") 
 ```
 
-# Discussion
+```{r}
+#| label: fig-session_year
+#| fig-cap: Comparison of Attendance between 2022 & 2024
+#| echo: false
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {.unnumbered}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
+cleaned_data |> 
+  ggplot(aes(x = session_year, fill = present)) +
+  geom_bar(position = "dodge2") 
 ```

### 2024-09-21 scripts/02-data_cleaning.R

@@ -15,15 +15,12 @@ cleaned_data <-
   raw_data |>
   janitor::clean_names() |>
   filter(committee == "City Council") |> 
-  #change from Y = present to 1 and N = absent to 0
-  mutate(present = ifelse(present == "Y", 1, 0)) |>
   #removing day from date 
   separate(col = session_date,
            into = c("session_year", "session_month"),
            sep = "-") |> 
   #selects columns of interest
-  select(first_name, last_name, committee, 
-         session_year, session_month, session_type, present)
+  select(committee, session_year, session_month, session_type, present)
 
 #### Save data ####
 write_csv(cleaned_data, "data/cleaned_data/cleaned_data.csv")

### 2024-09-21 scripts/03-test_data.R

@@ -10,8 +10,11 @@ raw_data <- read_csv("data/raw_data/unedited_data.csv")
 # Test for NAs
 all(is.na(raw_data))
 
+# Test for City Council Member
+raw_data$Committee == "City Council"
+
 #### Tests cleaned data ####
 cleaned_data <- read_csv("data/cleaned_data/cleaned_data.csv")
 
-# Test for NAs
+# Test for City Council Member
 cleaned_data$committee == "City Council"

### 2024-09-21 paper/paper.qmd
did some comment & cleaning of files + updated reference page + added a table counting the different types of sessions

@@
   - Wendy Yuan
-thanks: "Code and data are available at: LINK."
 date: "27 September 2024"
@@
 ```
+\newpage
 
@@
 # Data {#sec-data}
+
+```{r}
+#| label: fig-count
+#| fig-cap: Count of Number of Sessions per Year, Month, & Type
+#| echo: false
+
+table(cleaned_data$session_type)
+
+cleaned_data$session_month <- factor(cleaned_data$session_month, levels = month.abb)
+table(cleaned_data$session_month)
+
+table(cleaned_data$session_year)
+```
 
@@
 ggplot(cleaned_data, aes(x = session_type, fill = present)) +
-  geom_bar(position = "dodge2") 
+  geom_bar(position = "dodge2") + 
+  labs(
+    title = "Present vs Absence in Different Session Types",
+    x = "Session Type",
+    y = "Count"
+  )
 ```
@@
   ggplot(aes(x = session_month, fill = present)) +
-  geom_bar(position = "dodge2") 
+  geom_bar(position = "dodge2") + 
+  labs(
+    title = "Present vs Absence Over Different Months",
+    x = "Months",
+    y = "Count"
+  )
 ```
@@
 #| label: fig-session_year
-#| fig-cap: Comparison of Attendance between 2022 & 2024
+#| fig-cap: Comparison of Attendance between 2022 to 2024
 #| echo: false
@@
   ggplot(aes(x = session_year, fill = present)) +
-  geom_bar(position = "dodge2") 
+  geom_bar(position = "dodge") + 
+  labs(
+    title = "Present vs Absence Over Different Years",
+    x = "Years",
+    y = "Count"
+  )
 ```

### 2024-09-21 scripts/00-simulate_data.R

@@ -10,19 +10,22 @@ library(tidyverse)
 #### Simulate data ####
 set.seed(304)
 
+# creating inital variables
 observations <- 100
 year <- c(2022:2024)
 months <- c(1:12)
 type <- c("morning", "afternoon")
 
+# generating random samples of each variable of interest 
 session_type <- sample(type, observations, replace = TRUE)
 session_year <- sample(year, observations, replace = TRUE)
 session_month <- sample(months, observations, replace = TRUE)
 attendance <- rbinom(observations, 1, 0.75)
 
+# adding generated samples into a data frame
 sim_data <- 
   data.frame(session_type, session_year, session_month, attendance)
 
-#### Write_csv
+#### Writing csv ####
 write_csv(sim_data, file = "data/simulated_data/simulated_data.csv")
 

### 2024-09-21 scripts/01-download_data.R

@@ -5,7 +5,6 @@
 # Contact: w.yuan@mail.utoronto.ca
 
 #### Workspace setup ####
-
 library(opendatatoronto)
 library(dplyr)
 

### 2024-09-21 scripts/02-data_cleaning.R

@@ -7,6 +7,7 @@
 #### Workspace setup ####
 library(tidyverse)
 library(dplyr)
+library(lubridate)
 
 #### Clean data ####
 raw_data <- read_csv("data/raw_data/unedited_data.csv")
@@ -15,11 +16,15 @@ cleaned_data <-
   raw_data |>
   janitor::clean_names() |>
   filter(committee == "City Council") |> 
-  #removing day from date 
+  # changing present to be more readable
+  mutate(present = ifelse(present == "Y", "Yes", "No")) |>
+  # removing day from date 
   separate(col = session_date,
            into = c("session_year", "session_month"),
            sep = "-") |> 
-  #selects columns of interest
+  # changing months from numbers to words
+  mutate(session_month = month.abb[as.numeric(session_month)]) |>
+  # selects columns of interest
   select(committee, session_year, session_month, session_type, present)
 
 #### Save data ####

### 2024-09-21 paper/paper.qmd
wrote abstract & readme + uploaded initial sketches

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Investigating Toronto City Council Attendance"
+subtitle: "Determining the Ideal Meeting Types and Months for the 2022-2026 Term"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper looks at Toronto City Council attendance data and when meeting absences are highest and lowest.
+It specifically examines that evening meeting have greater absence rate compared to morning and afternoon.
+2022 showed very little absences while 2023 showed many missed attendances.
+Additionally, March, May, July, and October have are greater proportion of absences - showing a pattern that every couple months has less attendance."
 format: pdf
@@
 ```
+
 \newpage

### 2024-09-21 paper/paper.qmd
updated with the introduction

--- 2024-09-21 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Investigating Trends of Absences in Toronto City Council Attendance between 2022 to 2024"
+subtitle: "Determining the Ideal Meeting Types and Months for the 2022-2026 Term"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper looks at Toronto City Council attendance data and when meeting absences are highest and lowest.
+As Toronto City Council members are elected, it is expected that they attend meetings to voice the opinions of the population.
+Examining the data found that alternating months and evenings have greater absences.
+By identifying trends of absences, meetings can be better schedules to optimize efficiency and attendance."
 format: pdf
@@
 # Introduction
+
+In a democracy, individuals are expected to contribute their voices through voting.
+Likewise, elected officials should uphold their duty of representing the opinions of those who voted them in.
+So, the elected members of the Toronto City Council must attend their meetings to express their concerns to the population.
+
+This paper looks at the attendance of Toronto City Council members since the beginning of the 2022 term.
+The data used in this paper can be found on Open Data Toronto.
+First, this paper looks at the distribution of when most meetings are held to understand the data and biases better.
+Session type, month and year are examined to see if any trends can be avoided in the future to increase attendance.
+
+Session type refers to whether the session was held in the morning, afternoon, or evening.
+Session month shows which month the meeting was held likewise session year in which year the meeting happened.
+Finally, the present variable shows their attendance with yes if they were present and no if they were not.
+
+This paper found a trend of alternating months in absence, where one month will have high absence and the next will have low absence.
+Evening sessions also tended to have more absences.
+In 2022, there were very few absences while 2023 showed the opposite and 2024 has shown to stay in between for now.
+
+This paper is structured by beginning with the abstract and introduction.
+The remainder looks at how the data was cleaned, a table summarizing meeting types, and examines how each variable influences the attendance of City Council members.
 
@@
 
-\newpage
-
 # References

### 2024-09-22 paper/paper.qmd
finished writing data section - only missing references to do

@@
 
+To investigate the attendance of Toronto City Council members data is retrieved from Open Data Toronto.
+The raw data contained columns of Term, First Name, Last Name, Committee, MTG #, Session Date, Session Type, Session Start-End Time, Present.
+After importing the data into R, it was cleaned to make it more readable and contain only the necessary variables.
+Tests were also run on the data to make sure it is city council member data and no empty rows are there.
+
+First, the session date was made more readable by separating the year and month.
+Additionally, the month was changed to the word abbreviation form.
+The present variable was also made more readable by changing Y to Yes and N to No.
+Finally, the columns of interest, committee, session year, session month, session type, and present, were selected to be created into a new cleaned dataset.
+
+Between the different variables, type, month and year, they each have different frequency.
+Understanding the frequency, can help better understand the outcome of the data.
+Thus, three tables were made counting how many meetings were done in each circumstance.
+
 ```{r}
@@
 ```
+
+It can be seen that there are significantly less evening meetings while afternoon meetings are most common.
+Noticeably, January has no meetings likely because of external reasons.
+Meeting frequency of different months seem to change with no particular pattern.
+Finally, 2023 had the most meetings simply because it had the most time.
+
+First, to analysis which meeting type is best a bar chart is created comparing morning, afternoon, and evening times.
 
@@
 ```
+
+@fig-session_type has x-axis as the session type and y-axis as the count of how many times each meeting type occurred.
+The blue bars are the number of attendances while the red bars are the number of absences.
+It can be seen that afternoon has the most absences but it also has the most attendance.
+Notably, while evenings occurrences are low it still has a decent amount of absences.
+From the data, meetings would be better held during the afternoon or morning for the best attendance.
+
+Next, to find out which month had the best attendance rate a bar chart was graphed. The x-axis is the months and the y-axis is the count.
 
@@
 
+In @fig-session_month the blue bars is the count of attendance during the month and the red bars is the count of absences.
+There is a rough pattern where alternating months will have higher absences, such as March, May, and July.
+A pattern can not be determined for the attendance rather it reflect the frequency meetings were held according to months.
+In conclusion, finding out what causes the pattern of absence would help increase attendance in those months.
+
+Finally, a comparison between the years since the beginning of the term was done by plotting a bar chat.
+
 ```{r}
@@
 
+@fig-session_year has x-axis of years and y-axis counting attendance.
+The blue bars is the attendance of those who went to meetings while red bars counts those absent.
+Like how there were many meetings conducted in 2023 there are also many absence which makes sense.
+Therefore, it is likely that as 2024 continues the absences will increase showing similar pattern to 2023.
+Meanwhile, 2022 in this bar chart show little to no absent which may be due to external factors.
+Overall, investigating which meeting type, month and year gives an idea as to what causes absences during meetings but more variables should be looked at to get a bigger and better picture of the circumstances.
+
+\newpage
+
 # References
+
+TODO: figure out how to do references + in text references

### 2024-09-26 paper/paper.qmd
did some housekeeping cleaning + reuploaded new sketches

@@
 ---
-title: "Investigating Trends of Absences in Toronto City Council Attendance between 2022 to 2024"
+title: "Investigating Trends of Absences in Toronto City Council Attendance between 2022 to 2026"
 subtitle: "Determining the Ideal Meeting Types and Months for the 2022-2026 Term"
@@
   - Wendy Yuan
+thanks: "GitHub repo link: https://github.com/kiwindyy/CityCouncilAttendance"
 date: "27 September 2024"
@@
 cleaned_data <- read_csv(here::here("data/cleaned_data/cleaned_data.csv"))
+raw_data <- read_csv(here::here("data/raw_data/unedited_data.csv"))
 ```
-
-\newpage
 

### 2024-09-26 scripts/01-download_data.R
updated missing package in download data

@@ -7,6 +7,7 @@
 #### Workspace setup ####
 library(opendatatoronto)
 library(dplyr)
+library(readr)
 
 # get package
 package <- show_package("9426c94d-abdb-45fe-9a81-79be40d204d2")

### 2024-09-26 paper/paper.qmd
made more reproducible + updated abstract & introduction

@@
 date-format: long
-abstract: "This paper looks at Toronto City Council attendance data and when meeting absences are highest and lowest.
-As Toronto City Council members are elected, it is expected that they attend meetings to voice the opinions of the population.
-Examining the data found that alternating months and evenings have greater absences.
-By identifying trends of absences, meetings can be better schedules to optimize efficiency and attendance."
+abstract: "This paper looks at trends in absences for Toronto City Council members using data from Open Data Toronto (@ ).
+The data was used to observe which members had the most absences, absences trends over the years, and what meeting types had the most absences.
+It was found that evening meetings saw the most absences and absences are predicted to see an increase over the years.
+Through identifying trends of absences, we can help minimize the number of absences by ___ members which is important as when City Council members are elected, they are expected to attend meetings to voice the population’s opinions."
 format: pdf
@@
 library(dplyr)
+library(readr)
 
@@
 
-In a democracy, individuals are expected to contribute their voices through voting.
+The Canadian parliamentary system is a democracy, where the population is represented in government by a selection of elected officials.
+To maintain the democratic structure, individuals should engage in political participation to have their voices represented, primarily through voting.
 Likewise, elected officials should uphold their duty of representing the opinions of those who voted them in.
-So, the elected members of the Toronto City Council must attend their meetings to express their concerns to the population.
+Otherwise, failing to do so will cause a decrease in trust in the government.
+So, it is the job of the elected members of the Toronto City Council to attend their meetings and express the concerns of the population they represent.
 
-This paper looks at the attendance of Toronto City Council members since the beginning of the 2022 term.
-The data used in this paper can be found on Open Data Toronto.
-First, this paper looks at the distribution of when most meetings are held to understand the data and biases better.
-Session type, month and year are examined to see if any trends can be avoided in the future to increase attendance.
+This paper looks at the absences of Toronto City Council members since the beginning of the 2022 term.
+The data used in this paper is found on Open Data Toronto (@ ).
+The study examines which council members have the highest rates of absence, the fluctuation of absences over time, and which meeting type has the most absences.
+This paper works to fill the gap of focused analysis on absences patterns within the City Council which directly relates to the effectiveness of representation by council members.
 
-Session type refers to whether the session was held in the morning, afternoon, or evening.
-Session month shows which month the meeting was held likewise session year in which year the meeting happened.
-Finally, the present variable shows their attendance with yes if they were present and no if they were not.
+The data analysis focuses on examining patterns in the data for various causes of absences.
+It was found that evening meetings experienced the most absences and the trend over the years shows that absences are likely to increase in 2022.
+Additionally, ___ members were found to have the most absences.
+Rather than blaming individuals for absences, understanding these patterns allows for future targeted meeting scheduling to minimize absences.
+This ensures that elected officials are present to represent the voices of the general.
 
-This paper found a trend of alternating months in absence, where one month will have high absence and the next will have low absence.
-Evening sessions also tended to have more absences.
-In 2022, there were very few absences while 2023 showed the opposite and 2024 has shown to stay in between for now.
-
-This paper is structured by beginning with the abstract and introduction.
-The remainder looks at how the data was cleaned, a table summarizing meeting types, and examines how each variable influences the attendance of City Council members.
+The paper is structured as follows: section 2 discusses the raw and clean data used in the study and any limitations in the dataset.
+Section 3 shows graphical visualization of the analysis results using data from section 2.
+Finally, section 4 covers interpretations of each graph from section 3 and any weaknesses of this study.
 

### 2024-09-26 scripts/02-data_cleaning.R

@@ -8,6 +8,7 @@
 library(tidyverse)
 library(dplyr)
 library(lubridate)
+library(readr)
 
 #### Clean data ####
 raw_data <- read_csv("data/raw_data/unedited_data.csv")

### 2024-09-26 scripts/03-test_data.R

@@ -4,6 +4,8 @@
 # Date: 27 September 2024
 # Contact: w.yuan@mail.utoronto.ca
 
+library(readr)
+
 #### Tests raw data ####
 raw_data <- read_csv("data/raw_data/unedited_data.csv")
 

### 2024-09-26 paper/paper.qmd
updated data section + reorganized sections

@@
 format: pdf
+toc: true
 number-sections: true
@@
 library(readr)
+library(ggplot2)
 
@@
 The paper is structured as follows: section 2 discusses the raw and clean data used in the study and any limitations in the dataset.
-Section 3 shows graphical visualization of the analysis results using data from section 2.
+Section 3 shows a graphical visualization of the analysis results using data from section 2.
 Finally, section 4 covers interpretations of each graph from section 3 and any weaknesses of this study.
@@
 
-To investigate the attendance of Toronto City Council members data is retrieved from Open Data Toronto.
-The raw data contained columns of Term, First Name, Last Name, Committee, MTG #, Session Date, Session Type, Session Start-End Time, Present.
-After importing the data into R, it was cleaned to make it more readable and contain only the necessary variables.
-Tests were also run on the data to make sure it is city council member data and no empty rows are there.
+## Raw Data
 
-First, the session date was made more readable by separating the year and month.
-Additionally, the month was changed to the word abbreviation form.
-The present variable was also made more readable by changing Y to Yes and N to No.
-Finally, the columns of interest, committee, session year, session month, session type, and present, were selected to be created into a new cleaned dataset.
+The data used to investigate the attendance of Toronto City Council members is from the Open Data Toronto website (@ ).
+To measure how much City Council members worked to voice their thoughts - their attendance was examined.
+The attendance data is collected by sessions, with each new session beginning when a meeting is finished.
+A member was labelled as present if they attended any meeting sessions.
+Roll calls were performed to make sure the council member was present.
+The data is updated when the duration of the meeting have been published.
 
-Between the different variables, type, month and year, they each have different frequency.
-Understanding the frequency, can help better understand the outcome of the data.
-Thus, three tables were made counting how many meetings were done in each circumstance.
+The raw dataset features 10 variables and 6071 observations.
+@ tbl-raw-data shows the first 10 observations of the 10 variables: _id, Term, First Name, Last Name, Committee, MTG #, Session Date, Session Type, Session Start-End Time, and Present.
+‘_id’ is the number associated with each observation.
+‘Term’ shows which term the council members are a part of, which is 2022-2026 as this is the timeframe we are interested in.
+The ‘First name’ and ‘last name’ variables are the first and last names of each council member.
+‘Committee’ is the committee the council member is a part of and in this paper, we are only looking at the City Council.
+‘MTG #’ is the short form of meeting number.
+‘Session Date’ refers to the day the session was held, in year-month-date format.
+‘Session Type’ specifies whether the meeting took place during the Morning, Afternoon, or Evening.
+‘Session Start-End Time’ gives the exact time the session was held.
+Finally, ‘Present’ marks whether the City Council member was present or absent.
 
-```{r}
-#| label: fig-count
-#| fig-cap: Count of Number of Sessions per Year, Month, & Type
-#| echo: false
+## Cleaning Data
 
-table(cleaned_data$session_type)
+After importing the data into R (@ ) using the Open Data Toronto package (@ ), it is cleaned to keep only key variables of interest and make it easier for data analysis in Section 3.
+To clean the data and produce graphs with the data, the following packages in R (@ ) were used: tidyverse (@ ), dplyr (@ ), lubridate (@ ), knitr (@ ), and ggplot2 (@ ).
 
-cleaned_data$session_month <- factor(cleaned_data$session_month, levels = month.abb)
-table(cleaned_data$session_month)
+The key variables of interest are name, committee, session year, session type, and present.
+First, first and last names of City Council members are combined into 1 variable to make it easier to read in graphs.
+Then, the session date is coded to separate the year, month and date - keeping only the year.
+When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset.
+The cleaned dataset consists of 5 variables and 3061 observations as seen in @ tbl-cleaned-data.
+Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.
 
-table(cleaned_data$session_year)
-```
+## Data Limitations
 
-It can be seen that there are significantly less evening meetings while afternoon meetings are most common.
-Noticeably, January has no meetings likely because of external reasons.
-Meeting frequency of different months seem to change with no particular pattern.
-Finally, 2023 had the most meetings simply because it had the most time.
+There were no similar datasets containing information about the attendance of Toronto City Council members on Open Data Toronto.
+If this paper were to be considered with different variables measuring the effectiveness of City Council members’ representation of the population then other datasets could be used.
+An example of a dataset that could have been an alternative is “City Council and Committees Meeting Schedule Reports” (@ ) which looks at the meeting schedule reports.
+The use of other datasets would have changed the research of this paper.
 
-First, to analysis which meeting type is best a bar chart is created comparing morning, afternoon, and evening times.
+A limitation faced by this dataset is that some sessions are closed sessions.
+Closed sessions are those “closed to the public under sections 27-38 of the Council Procedures by law” (@ ).
+These closed sessions could be viewed as null observations potentially altering the results of data analysis.
+Another limitation of this dataset is the lack of information on the reason for absence.
+City Council members could have valid excuses such as illness or urgent emergencies.
+Without this information, the conclusions lack the reasoning behind the results.
+
+# Results
+
+## Toronto City Council member absence
 
@@
 
-@fig-session_type has x-axis as the session type and y-axis as the count of how many times each meeting type occurred.
-The blue bars are the number of attendances while the red bars are the number of absences.
-It can be seen that afternoon has the most absences but it also has the most attendance.
-Notably, while evenings occurrences are low it still has a decent amount of absences.
-From the data, meetings would be better held during the afternoon or morning for the best attendance.
-
-Next, to find out which month had the best attendance rate a bar chart was graphed. The x-axis is the months and the y-axis is the count.
+## Toronto City Council absence over time since 2022
 
@@
 
-In @fig-session_month the blue bars is the count of attendance during the month and the red bars is the count of absences.
-There is a rough pattern where alternating months will have higher absences, such as March, May, and July.
-A pattern can not be determined for the attendance rather it reflect the frequency meetings were held according to months.
-In conclusion, finding out what causes the pattern of absence would help increase attendance in those months.
-
-Finally, a comparison between the years since the beginning of the term was done by plotting a bar chat.
+## Toronto City Council meeting type absence
 
@@
 
-@fig-session_year has x-axis of years and y-axis counting attendance.
-The blue bars is the attendance of those who went to meetings while red bars counts those absent.
-Like how there were many meetings conducted in 2023 there are also many absence which makes sense.
-Therefore, it is likely that as 2024 continues the absences will increase showing similar pattern to 2023.
-Meanwhile, 2022 in this bar chart show little to no absent which may be due to external factors.
-Overall, investigating which meeting type, month and year gives an idea as to what causes absences during meetings but more variables should be looked at to get a bigger and better picture of the circumstances.
+# Discussion
+
+## Weaknesses and next steps
+
+\newpage
+
+\appendix
+
+# Appendix {-}
+
+## Raw Data
+
+## Cleaned Data
 
@@
 # References
-
-TODO: figure out how to do references + in text references

### 2024-09-26 paper/paper.qmd
Updated data results section

@@
 The paper is structured as follows: section 2 discusses the raw and clean data used in the study and any limitations in the dataset.
-Section 3 shows a graphical visualization of the analysis results using data from section 2.
-Finally, section 4 covers interpretations of each graph from section 3 and any weaknesses of this study.
+Section 3 shows a graphical visualization of the data from section 2 and explains some trends that can be seen.
+Finally, section 4 covers interpretations of the data from section 3 and any weaknesses of this study.
 
@@
 
-## Toronto City Council member absence
+## Toronto City Council member's absence
+
+Graphs can help better understand the different relationships between each variable and the number of absences in the Toronto City Council.
+@ fig-session-member shows the number of absences by each Toronto City Council member in a bar graph.
+The graph’s x-axis represents the different members of the City Council and the y-axis is counting the number of absences.
 
 ```{r}
-#| label: fig-session_type
+#| label: fig-session-member
 #| fig-cap: Comparison of Attendance in Morning, Afternoon, vs Evening Session
@@
 
+In @ fig-session-member the taller the bar is the more absences that City Council member has had while the opposite is true, the lower the bar the fewer absences.
+This graph helps visualize the number of absences, showing no clear pattern or trend between the members.
+It can be seen that some members rarely miss sessions such as ___ while others, such as ___, have missed notably more.
+Thus, @ fig-session-member shows a general idea of individual absences, this graph makes it difficult to follow trends in the future.
+
 ## Toronto City Council absence over time since 2022
 
+To further our understanding of the data, @ fig-session-year works as a comparison between the years since the beginning of the term and member absences.
+The x-axis is the years since the beginning of the term and the y-axis counts the number of absences.
+
 ```{r}
-#| label: fig-session_month
+#| label: fig-session-year
 #| fig-cap: Comparison of Attendance over Different Months
@@
 
+The bar of 2022 in @ fig-session-year is very small meaning there were few absences during sessions in this period.
+It is likely due to a combination of factors but the most likely remains the accessibility to virtual meetings during the pandemic.
+The second bar, 2023, in the @ fig-session-year is significantly greater than 2022.
+Noticing the increasing trend, we predict a likely increase in absences.
+Interpreting the 2024 bar, it is shorter than 2023 but still greater than 2022.
+Knowing months are remaining in the year, it is likely that the absences in 2024 will increase, showing similar patterns to 2023.
+
 ## Toronto City Council meeting type absence
 
+The type of session may also influence whether the City Council member is present or absent.
+@ fig-session-type shows the distribution of absences across different session types - morning, afternoon, and evening - in the form of a bar graph.
+A bar graph is used since it displays the counts of categorical data.
+The x-axis is the different session types and the y-axis is the number of absences for each category.
+@ fig-session-type shows that afternoon sessions have the highest number of absences, followed by morning sessions, while evening sessions show relatively fewer.
+By visualizing the data, we can examine patterns and trends that might go unnoticed and the reasons behind them.
+
 ```{r}
-#| label: fig-session_year
+#| label: fig-session-type
 #| fig-cap: Comparison of Attendance between 2022 to 2024

### 2024-09-26 paper/paper.qmd
updated: README - new llm usage paper.qmd - finished writing discussion & made new graphs data-cleaning.R - changed cleaned data code

@@
 date-format: long
-abstract: "This paper looks at trends in absences for Toronto City Council members using data from Open Data Toronto (@ ).
+abstract: "This paper looks at trends in absences for Toronto City Council members using data from Open Data Toronto (@toronto-city-council-attendance).
 The data was used to observe which members had the most absences, absences trends over the years, and what meeting types had the most absences.
@@
 This paper looks at the absences of Toronto City Council members since the beginning of the 2022 term.
-The data used in this paper is found on Open Data Toronto (@ ).
+The data used in this paper is found on Open Data Toronto (@toronto-city-council-attendance).
 The study examines which council members have the highest rates of absence, the fluctuation of absences over time, and which meeting type has the most absences.
@@
 
-The data used to investigate the attendance of Toronto City Council members is from the Open Data Toronto website (@ ).
+The data used to investigate the attendance of Toronto City Council members is from the Open Data Toronto website (@toronto-city-council-attendance).
 To measure how much City Council members worked to voice their thoughts - their attendance was examined.
@@
 
-The raw dataset features 10 variables and 6071 observations.
-@ tbl-raw-data shows the first 10 observations of the 10 variables: _id, Term, First Name, Last Name, Committee, MTG #, Session Date, Session Type, Session Start-End Time, and Present.
+```{r echo=FALSE}
+#| label: tbl-raw-data
+head(raw_data, 10)
+
+```
+
+The raw dataset features `r ncol(raw_data)` variables and `r nrow(raw_data)` observations.
+@tbl-raw-data shows the first 10 observations of the `r ncol(raw_data)` variables: _id, Term, First Name, Last Name, Committee, MTG #, Session Date, Session Type, Session Start-End Time, and Present.
 ‘_id’ is the number associated with each observation.
@@
 
-After importing the data into R (@ ) using the Open Data Toronto package (@ ), it is cleaned to keep only key variables of interest and make it easier for data analysis in Section 3.
-To clean the data and produce graphs with the data, the following packages in R (@ ) were used: tidyverse (@ ), dplyr (@ ), lubridate (@ ), knitr (@ ), and ggplot2 (@ ).
+After importing the data into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in Section 3.
+To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr) and ggplot2 (@ggplot2).
 
+```{r echo=FALSE}
+#| label: tbl-cleaned-data
+head(cleaned_data, 10)
+
+```
 The key variables of interest are name, committee, session year, session type, and present.
-First, first and last names of City Council members are combined into 1 variable to make it easier to read in graphs.
-Then, the session date is coded to separate the year, month and date - keeping only the year.
+First and last names of City Council members are combined into 1 variable to make it easier to read in graphs.
+Then, the session date is decoded to separate the year, month and date - keeping only the year.
 When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset.
-The cleaned dataset consists of 5 variables and 3061 observations as seen in @ tbl-cleaned-data.
+The cleaned dataset consists of 5 variables as seen in @tbl-cleaned-data.
 Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.
@@
 If this paper were to be considered with different variables measuring the effectiveness of City Council members’ representation of the population then other datasets could be used.
-An example of a dataset that could have been an alternative is “City Council and Committees Meeting Schedule Reports” (@ ) which looks at the meeting schedule reports.
+An example of a dataset that could have been an alternative is “City Council and Committees Meeting Schedule Reports” (@toronto-council-attendance) which looks at the meeting schedule reports.
 The use of other datasets would have changed the research of this paper.
@@
 A limitation faced by this dataset is that some sessions are closed sessions.
-Closed sessions are those “closed to the public under sections 27-38 of the Council Procedures by law” (@ ).
+Closed sessions are those “closed to the public under sections 27-38 of the Council Procedures by law” (@toronto-city-council-attendance).
 These closed sessions could be viewed as null observations potentially altering the results of data analysis.
@@
 
-Graphs can help better understand the different relationships between each variable and the number of absences in the Toronto City Council.
-@ fig-session-member shows the number of absences by each Toronto City Council member in a bar graph.
+Graphs can help better understand the relationships between each variable and the number of absences in the Toronto City Council.
+@fig-session-member shows the number of absences by each Toronto City Council member in a bar graph.
 The graph’s x-axis represents the different members of the City Council and the y-axis is counting the number of absences.
@@
 #| label: fig-session-member
-#| fig-cap: Comparison of Attendance in Morning, Afternoon, vs Evening Session
+#| fig-cap: Comparision of Absences by name
 #| echo: false
 
-ggplot(cleaned_data, aes(x = session_type, fill = present)) +
-  geom_bar(position = "dodge2") + 
+# Filter the data for absences (where 'present' is 'No') and count them by name
+absence_by_member <- cleaned_data %>%
+  filter(present == "No") %>%
+  group_by(name) %>%
+  summarise(absence_count = n())
+
+ggplot(absence_by_member, aes(x = reorder(name, -absence_count), y = absence_count)) +
+  geom_bar(stat = "identity", fill = "cyan3") +
   labs(
-    title = "Present vs Absence in Different Session Types",
-    x = "Session Type",
-    y = "Count"
-  )
+    title = "Number of Absences by Each Member",
+    x = "Council Member",
+    y = "Number of Absences"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
 
-In @ fig-session-member the taller the bar is the more absences that City Council member has had while the opposite is true, the lower the bar the fewer absences.
-This graph helps visualize the number of absences, showing no clear pattern or trend between the members.
+In @fig-session-member the taller the bar is the more absences that City Council member has had while the opposite is true, the lower the bar the fewer absences.
 It can be seen that some members rarely miss sessions such as ___ while others, such as ___, have missed notably more.
-Thus, @ fig-session-member shows a general idea of individual absences, this graph makes it difficult to follow trends in the future.
+Thus, @fig-session-member shows a general idea of individual absences, this graph makes it difficult to follow trends in the future.
 
@@
 
-To further our understanding of the data, @ fig-session-year works as a comparison between the years since the beginning of the term and member absences.
+To further our understanding of the data, @fig-session-year compares the number of member absences over the years since the beginning of the term.
 The x-axis is the years since the beginning of the term and the y-axis counts the number of absences.
@@
 #| label: fig-session-year
-#| fig-cap: Comparison of Attendance over Different Months
+#| fig-cap: Comparison of Member Absences by Year
 #| echo: false
 
-cleaned_data |> 
-  ggplot(aes(x = session_month, fill = present)) +
-  geom_bar(position = "dodge2") + 
+# Filter the data to only include rows where members were absent (present == "No")
+absent_data <- cleaned_data %>%
+  filter(present == "No") %>%
+  group_by(session_year) %>%
+  summarise(absence_count = n())  # Count the number of absences by year
+
+# Create the plot
+ggplot(absent_data, aes(x = session_year, y = absence_count)) +
+  geom_bar(stat = "identity", fill = "cyan3") +  # Use a bar plot and custom color
   labs(
-    title = "Present vs Absence Over Different Months",
-    x = "Months",
-    y = "Count"
-  )
+    title = "# absence over years 2022-2024",
+    x = "Year",
+    y = "# Absences"
+  ) +
+  theme_minimal()
 ```
 
-The bar of 2022 in @ fig-session-year is very small meaning there were few absences during sessions in this period.
+The bar of 2022 in @fig-session-year is small meaning there were few absences during sessions in this period.
 It is likely due to a combination of factors but the most likely remains the accessibility to virtual meetings during the pandemic.
-The second bar, 2023, in the @ fig-session-year is significantly greater than 2022.
+The second bar, 2023, in the @fig-session-year is significantly greater than 2022.
 Noticing the increasing trend, we predict a likely increase in absences.
@@
 The type of session may also influence whether the City Council member is present or absent.
-@ fig-session-type shows the distribution of absences across different session types - morning, afternoon, and evening - in the form of a bar graph.
+@fig-session-type shows the distribution of absences across different session types - morning, afternoon, and evening - in the form of a bar graph.
 A bar graph is used since it displays the counts of categorical data.
 The x-axis is the different session types and the y-axis is the number of absences for each category.
-@ fig-session-type shows that afternoon sessions have the highest number of absences, followed by morning sessions, while evening sessions show relatively fewer.
+@fig-session-type shows that afternoon sessions have the most absences, followed by morning sessions, while evening sessions show relatively fewer.
 By visualizing the data, we can examine patterns and trends that might go unnoticed and the reasons behind them.
@@
 #| label: fig-session-type
-#| fig-cap: Comparison of Attendance between 2022 to 2024
+#| fig-cap: Comparison of Attendance by session type between 2022 to 2024
 #| echo: false
 
-cleaned_data |> 
-  ggplot(aes(x = session_year, fill = present)) +
-  geom_bar(position = "dodge") + 
+# Filter the data for absences (present == "No") and group by session type
+absence_data <- cleaned_data %>%
+  filter(present == "No") %>%
+  group_by(session_type) %>%
+  summarise(absence_count = n())  # Count the number of absences for each session type
+
+# Create the plot
+ggplot(absence_data, aes(x = session_type, y = absence_count)) +
+  geom_bar(stat = "identity", fill = "cyan3") +  # Use horizontal bar plot and a custom color
   labs(
-    title = "Present vs Absence Over Different Years",
-    x = "Years",
-    y = "Count"
-  )
+    title = "# absence of different mettings types",
+    x = "Type",
+    y = "# Absences"
+  ) +
+  theme_minimal()
 ```
@@
 
+@fig-session-member from Section 3 laid out the information about the absences of each Toronto City Council member.
+It can be seen that __ has the highest absences, while __ has the lowest.
+The variation between members could be problematic especially when the council member fails to represent the population.
+Additionally, it could suggest differences in engagement, personal schedules, or emergency circumstances.
+Determining those with the most absences allows steps to be taken to help better accommodate their schedules.
+
+@fig-session-type shows that afternoon meetings have more absences than morning or evening ones.
+The timing of the meetings could be a factor contributing to the absences of council members.
+This pattern could mean that members are facing conflicting schedules.
+Morning sessions show a significant decrease in absences so the earlier timing could allow better attendance.
+Finally, evening sessions saw the lowest absence rate possibly providing an alternative meeting time.
+Understanding the effects of session timing on attendance would lead to better and more engaging discussions about policies as more members are present to voice their opinions.
+
+Finally, @fig-session-year looks at the trend of absences over the year since the beginning of the term.
+There is an increase in absences from 2022-2023.
+This could be linked to pandemic-related transitions from virtual to in-person, as virtual meetings are more accessible.
+2024 saw fewer absences potentially signalling post-pandemic adjustments and adapting to new conditions.
+Absences in 2024 could still increase in the remaining months due to session changes, holidays, and other disruptions.
+Overall, the data shows trends in absences for individual members and over time.
+Factors such as post-pandemic adjustment, scheduling conflicts and external commitments all likely contribute to the observed patterns.
+Further
+
 ## Weaknesses and next steps
 
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-## Raw Data
-
-## Cleaned Data
+A key limitation of this paper is the limited scope of analysis.
+There is a focus on the number of absences without studying potential underlying causes.
+Many hidden factors, such as health and personal emergencies, likely contribute to these absences.
+Additionally, there lacks a deeper examination of variables such as council member’s workload, length of session and other social factors.
+Without these considerations, there are gaps in understanding the broader context of the absences.
+The next steps to improve the analysis of City Council member absence would include further investigations into the reasons behind individual absences and exploring more strategies to improve participation in future sessions.
 

### 2024-09-26 scripts/02-data_cleaning.R

@@ -19,14 +19,12 @@ cleaned_data <-
   filter(committee == "City Council") |> 
   # changing present to be more readable
   mutate(present = ifelse(present == "Y", "Yes", "No")) |>
-  # removing day from date 
-  separate(col = session_date,
-           into = c("session_year", "session_month"),
-           sep = "-") |> 
-  # changing months from numbers to words
-  mutate(session_month = month.abb[as.numeric(session_month)]) |>
+  mutate(name = paste(first_name, last_name, sep = " ")) |>
+  # separate year from session_date
+  mutate(session_year = substr(session_date, 1, 4)) |> 
   # selects columns of interest
-  select(committee, session_year, session_month, session_type, present)
+  select(name, committee, session_year, session_type, present)
 
 #### Save data ####
 write_csv(cleaned_data, "data/cleaned_data/cleaned_data.csv")
+

### 2024-09-26 paper/paper.qmd
Updated llm usage chat history & filled in any remaining gaps in the report

@@
 It was found that evening meetings saw the most absences and absences are predicted to see an increase over the years.
-Through identifying trends of absences, we can help minimize the number of absences by ___ members which is important as when City Council members are elected, they are expected to attend meetings to voice the population’s opinions."
+Through identifying trends of absences, we can help minimize the number of absences which is important as when City Council members are elected, they are expected to attend meetings to voice the population’s opinions."
 format: pdf
@@
 It was found that evening meetings experienced the most absences and the trend over the years shows that absences are likely to increase in 2022.
-Additionally, ___ members were found to have the most absences.
-Rather than blaming individuals for absences, understanding these patterns allows for future targeted meeting scheduling to minimize absences.
+Additionally, Michael Thompson, Jaye Robinson, and Vincent Crisanti were found to have the most absences Rather than blaming individuals for absences, understanding these patterns allows for future targeted meeting scheduling to minimize absences.
 This ensures that elected officials are present to represent the voices of the general.
@@
 To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr) and ggplot2 (@ggplot2).
+ChatGPT (@ChatGPT) was used to help the coding of cleaning, testing, and graphing data.
 
@@
 In @fig-session-member the taller the bar is the more absences that City Council member has had while the opposite is true, the lower the bar the fewer absences.
-It can be seen that some members rarely miss sessions such as ___ while others, such as ___, have missed notably more.
+It can be seen that some members rarely miss sessions such as Amber Morley, Olivia Chow, and Paula Fletcher.
+While Michael Thompson, Jaye Robinson, and Vincent Crisanti have missed many meetings.
 Thus, @fig-session-member shows a general idea of individual absences, this graph makes it difficult to follow trends in the future.
@@
 @fig-session-member from Section 3 laid out the information about the absences of each Toronto City Council member.
-It can be seen that __ has the highest absences, while __ has the lowest.
+It can be seen that Michael Thompson has the highest absences, followed by Jaye Robinson and Vincent Crisanti.
+While Amber Morley, Olivia Chow, and Paula Fletcher have missed few meetings.
 The variation between members could be problematic especially when the council member fails to represent the population.

### 2024-09-26 paper/paper.qmd
Updated LLM usage & some formating in paper

@@
 library(ggplot2)
+library(knitr)
 
@@
 
-# Introduction
+# Introduction {#sec-intro}
 
@@
 
-The paper is structured as follows: section 2 discusses the raw and clean data used in the study and any limitations in the dataset.
-Section 3 shows a graphical visualization of the data from section 2 and explains some trends that can be seen.
-Finally, section 4 covers interpretations of the data from section 3 and any weaknesses of this study.
+The paper is structured as follows: [@sec-data] discusses the raw and clean data used in the study and any limitations in the dataset.
+[@sec-results] shows a graphical visualization of the data from [@sec-data] and explains some trends that can be seen.
+Finally, [@sec-discussion] covers interpretations of the data from [@sec-results] and any weaknesses of this study.
 
@@
 #| label: tbl-raw-data
-head(raw_data, 10)
+#| tbl-cap: Raw Data of Toronto City Council Members' Attendance
 
+# generate first 10 observations of raw data
+kable(head(raw_data, 10)) 
 ```
@@
 
-After importing the data into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in Section 3.
+After importing the data into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in [@sec-results].
 To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr) and ggplot2 (@ggplot2).
@@
 #| label: tbl-cleaned-data
-head(cleaned_data, 10)
+#| tbl-cap: Cleaned Data of Toronto City Council Members' Attendance
 
+# generate first 10 observations of cleaned data
+kable(head(raw_data, 10)) 
 ```
@@
 
-# Results
+# Results {#sec-results}
 
@@
 #| label: fig-session-member
-#| fig-cap: Comparision of Absences by name
+#| fig-cap: Comparision of Absences by City Council Member
 #| echo: false
@@
 
+# generate graph for number of absences by each member
 ggplot(absence_by_member, aes(x = reorder(name, -absence_count), y = absence_count)) +
@@
 To further our understanding of the data, @fig-session-year compares the number of member absences over the years since the beginning of the term.
-The x-axis is the years since the beginning of the term and the y-axis counts the number of absences.
+The x-axis is the years since the beginning of the term until the present year and the y-axis counts the number of absences.
 
@@
 #| label: fig-session-year
-#| fig-cap: Comparison of Member Absences by Year
+#| fig-cap: Comparison of Absences by Year
 #| echo: false
@@
   labs(
-    title = "# absence over years 2022-2024",
+    title = "Number of Absence Over Years 2022-2024",
     x = "Year",
-    y = "# Absences"
+    y = "Number of Absences"
   ) +
@@
 #| label: fig-session-type
-#| fig-cap: Comparison of Attendance by session type between 2022 to 2024
+#| fig-cap: Comparison of Absences by Session Type
 #| echo: false
@@
   labs(
-    title = "# absence of different mettings types",
-    x = "Type",
-    y = "# Absences"
+    title = "Number Absence for Different Meetings Types",
+    x = "Session Type",
+    y = "Number of Absences"
   ) +
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
-@fig-session-member from Section 3 laid out the information about the absences of each Toronto City Council member.
+@fig-session-member from [@sec-results] laid out the information about the absences of each Toronto City Council member.
 It can be seen that Michael Thompson has the highest absences, followed by Jaye Robinson and Vincent Crisanti.

### 2024-09-27 paper/paper.qmd
updated gitignore, simulated data, data tests

@@
 ---
-title: "Investigating Trends of Absences in Toronto City Council Attendance between 2022 to 2026"
-subtitle: "Determining the Ideal Meeting Types and Months for the 2022-2026 Term"
+title: "Maximizing Council Participation: A Study on Toronto City Council Absences (2022-2026)"
 author: 
   - Wendy Yuan
-thanks: "GitHub repo link: https://github.com/kiwindyy/CityCouncilAttendance"
+thanks: "Code and data supporting this analysis is available at: https://github.com/kiwindyy/CityCouncilAttendance"
 date: "27 September 2024"
@@
 After importing the data into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in [@sec-results].
-To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr) and ggplot2 (@ggplot2).
+To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr), randomNames (@randomNames) and ggplot2 (@ggplot2).
 ChatGPT (@ChatGPT) was used to help the coding of cleaning, testing, and graphing data.

### 2024-09-27 scripts/00-simulate_data.R

@@ -6,26 +6,54 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(randomNames)
+library(lubridate)
 
 #### Simulate data ####
-set.seed(304)
+set.seed(1009122423)
+
+# Generate random start and end times
+generate_random_time_range <- function() {
+  start_time <- sprintf("%02d:%02d%s", sample(1:12, 1), sample(0:59, 1), sample(c("AM", "PM"), 1))
+  end_time <- sprintf("%02d:%02d%s", sample(1:12, 1), sample(0:59, 1), sample(c("AM", "PM"), 1))
+  time_range <- paste(start_time, "-", end_time)
+  
+  return(time_range)
+}
 
 # creating inital variables
+begin_year = 2022
+end_year = 2026
 observations <- 100
-year <- c(2022:2024)
-months <- c(1:12)
-type <- c("morning", "afternoon")
+names <- randomNames(10)
+committee_names <- c("City Council", "CreateTO", "Executive Committee", "Toronto Parking Authority", "Audit Committee")
+session_dates <- as.Date(sample(seq.Date(from = as.Date("2022-01-01"), to = as.Date("2024-12-31"), by = "day"), 20, replace = FALSE))
+mtgs <- c(1:4)
+type <- c("Morning", "Afternoon", "Evening")
+meeting_times <- replicate(10, generate_random_time_range())
 
-# generating random samples of each variable of interest 
-session_type <- sample(type, observations, replace = TRUE)
-session_year <- sample(year, observations, replace = TRUE)
-session_month <- sample(months, observations, replace = TRUE)
-attendance <- rbinom(observations, 1, 0.75)
+# generating random samples of each variable of interest
+full_names <- sample(names, observations, replace = TRUE)
+split_result <- strsplit(full_names, ", ")
+split_result <- unlist(split_result)
+`First Name` <- split_result[seq(2, length(split_result), by = 2)]
+`Last Name` <- split_result[seq(1, length(split_result), by = 2)]
+Committee <- sample(committee_names, observations, replace = TRUE)
+`MTG #` <- sample(mtgs, observations, replace=TRUE)
+session_type <- sample(type, observations, prob=c(0.4, 0.4, 0.2), replace = TRUE)
+session_date <- sample(session_dates, observations, replace = TRUE)
+session_time <- sample(meeting_times, observations, replace = TRUE)
[…]

### 2024-09-27 scripts/01-download_data.R

@@ -25,5 +25,3 @@ data
 
 #### Save data ####
 write_csv(data, "data/raw_data/unedited_data.csv") 
-
-         

### 2024-09-27 scripts/03-test_data.R

@@ -6,17 +6,33 @@
 
 library(readr)
 
+# creating inital variables
+begin_year = 2022
+end_year = 2026
+
 #### Tests raw data ####
 raw_data <- read_csv("data/raw_data/unedited_data.csv")
 
 # Test for NAs
 all(is.na(raw_data))
 
-# Test for City Council Member
-raw_data$Committee == "City Council"
-
 #### Tests cleaned data ####
 cleaned_data <- read_csv("data/cleaned_data/cleaned_data.csv")
 
+#Test first last name deliminator
+all(sapply(cleaned_data$name, function(name) {
+  split_name <- strsplit(name, " ")[[1]]  # Split by space
+  length(split_name) == 2                 # Check if it splits into exactly 2 parts
+}))
+
 # Test for City Council Member
-cleaned_data$committee == "City Council"
+all(cleaned_data$committee == "City Council")
+
+# Test session year
+all(cleaned_data$session_year >= begin_year & cleaned_data$session_year <= end_year & !is.na(cleaned_data$session_year))
+
+#Test for Session types
+all(cleaned_data$session_type %in% c("Morning", "Afternoon", "Evening"))
+
+#Test for present Yes or No
+all(cleaned_data$present %in% c("Yes", "No") & !is.na(cleaned_data$present))

### 2024-09-27 paper/paper.qmd
final small writing changes

@@
 Otherwise, failing to do so will cause a decrease in trust in the government.
+There are 3 levels of governments in Canada: federal, provincial, and municipal.
+The municipal government works on a city-wide scale, directly working to address local concerns.
 So, it is the job of the elected members of the Toronto City Council to attend their meetings and express the concerns of the population they represent.
@@
 [@sec-results] shows a graphical visualization of the data from [@sec-data] and explains some trends that can be seen.
-Finally, [@sec-discussion] covers interpretations of the data from [@sec-results] and any weaknesses of this study.
+Next, [@sec-discussion] covers interpretations of the data from [@sec-results] and any weaknesses of this study.
+Finally, [@sec-appendix] includes additional supporting data and [@sec-references] contains the references used in this paper.
 
@@
 
-## Cleaning Data
-
-After importing the data into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in [@sec-results].
-To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr), randomNames (@randomNames) and ggplot2 (@ggplot2).
+The data was into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in [@sec-results].
+The key variables of interest are name, committee, session year, session type, and present.
+@tbl-cleaned-data shows first 10 observations of the cleaned data and [@sec-appendix] explains in more detail how the data was cleaned.
+The following packages in R (@citeR) were used in the data analysis process: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr), randomNames (@randomNames) and ggplot2 (@ggplot2).
 ChatGPT (@ChatGPT) was used to help the coding of cleaning, testing, and graphing data.
-
-```{r echo=FALSE}
-#| label: tbl-cleaned-data
-#| tbl-cap: Cleaned Data of Toronto City Council Members' Attendance
-
-# generate first 10 observations of cleaned data
-kable(head(raw_data, 10)) 
-```
-The key variables of interest are name, committee, session year, session type, and present.
-First and last names of City Council members are combined into 1 variable to make it easier to read in graphs.
-Then, the session date is decoded to separate the year, month and date - keeping only the year.
-When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset.
-The cleaned dataset consists of 5 variables as seen in @tbl-cleaned-data.
-Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.
 
@@
 
-# References
+# Appendix {#sec-appendix}
+
+## Cleaning Data
+
+```{r echo=FALSE}
+#| label: tbl-cleaned-data
+#| tbl-cap: Cleaned Data of Toronto City Council Members' Attendance
+
+# generate first 10 observations of cleaned data
+kable(head(raw_data, 10)) 
+```
+
+First and last names of City Council members are combined into 1 variable to make it easier to read in graphs.
+Then, the session date is decoded to separate the year, month and date - keeping only the year.
+When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset.
+The cleaned dataset consists of 5 variables as seen in @tbl-cleaned-data.
+Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.
+
+\newpage
+
+# References {#sec-references}

### 2024-09-27 paper/paper.qmd
page styling changes
fixing lintr style

@@
 # generate graph for number of absences by each member
-ggplot(absence_by_member, aes(x = reorder(name, -absence_count), y = absence_count)) +
+ggplot(absence_by_member, aes(x = reorder(name, -absence_count),
+                              y = absence_count)) +
   geom_bar(stat = "identity", fill = "cyan3") +
@@
 
-# Filter the data to only include rows where members were absent (present == "No")
+# Filter data to only include rows where members were absent (present == "No")
 absent_data <- cleaned_data %>%
@@
 ggplot(absent_data, aes(x = session_year, y = absence_count)) +
-  geom_bar(stat = "identity", fill = "cyan3") +  # Use a bar plot and custom color
+  # Use a bar plot and custom color
+  geom_bar(stat = "identity", fill = "cyan3") +
   labs(
@@
   group_by(session_type) %>%
-  summarise(absence_count = n())  # Count the number of absences for each session type
+  # Count the number of absences for each session type
+  summarise(absence_count = n())
 
@@
 ggplot(absence_data, aes(x = session_type, y = absence_count)) +
-  geom_bar(stat = "identity", fill = "cyan3") +  # Use horizontal bar plot and a custom color
+  # Use horizontal bar plot and a custom color
+  geom_bar(stat = "identity", fill = "cyan3") +
   labs(
@@
 ```
+
+\newpage
 

### 2024-09-27 scripts/00-simulate_data.R
additional lintr edits

@@ -14,20 +14,25 @@ set.seed(1009122423)
 
 # Generate random start and end times
 generate_random_time_range <- function() {
-  start_time <- sprintf("%02d:%02d%s", sample(1:12, 1), sample(0:59, 1), sample(c("AM", "PM"), 1))
-  end_time <- sprintf("%02d:%02d%s", sample(1:12, 1), sample(0:59, 1), sample(c("AM", "PM"), 1))
+  start_time <- sprintf("%02d:%02d%s", sample(1:12, 1), sample(0:59, 1),
+                        sample(c("AM", "PM"), 1))
+  end_time <- sprintf("%02d:%02d%s", sample(1:12, 1), sample(0:59, 1),
+                      sample(c("AM", "PM"), 1))
   time_range <- paste(start_time, "-", end_time)
-  
   return(time_range)
 }
 
 # creating inital variables
-begin_year = 2022
-end_year = 2026
+begin_year <- 2022
+end_year <- 2026
 observations <- 100
 names <- randomNames(10)
-committee_names <- c("City Council", "CreateTO", "Executive Committee", "Toronto Parking Authority", "Audit Committee")
-session_dates <- as.Date(sample(seq.Date(from = as.Date("2022-01-01"), to = as.Date("2024-12-31"), by = "day"), 20, replace = FALSE))
+committee_names <- c("City Council", "CreateTO", "Executive Committee",
+                     "Toronto Parking Authority", "Audit Committee")
+session_dates <- as.Date(sample(seq.Date(from = as.Date("2022-01-01"),
+                                         to = as.Date("2024-12-31"),
+                                         by = "day"),
+                                20, replace = FALSE))
 mtgs <- c(1:4)
 type <- c("Morning", "Afternoon", "Evening")
 meeting_times <- replicate(10, generate_random_time_range())
@@ -36,19 +41,24 @@ meeting_times <- replicate(10, generate_random_time_range())
 full_names <- sample(names, observations, replace = TRUE)
 split_result <- strsplit(full_names, ", ")
 split_result <- unlist(split_result)
-`First Name` <- split_result[seq(2, length(split_result), by = 2)]
-`Last Name` <- split_result[seq(1, length(split_result), by = 2)]
-Committee <- sample(committee_names, observations, replace = TRUE)
-`MTG #` <- sample(mtgs, observations, replace=TRUE)
-session_type <- sample(type, observations, prob=c(0.4, 0.4, 0.2), replace = TRUE)
[…]

### 2024-09-27 scripts/01-download_data.R

@@ -16,12 +16,14 @@ package
 # get all resources for this package
 resources <- list_package_resources("9426c94d-abdb-45fe-9a81-79be40d204d2")
 
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+# identify datastore resources; by default, Toronto Open Data sets datastore
+# resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in%
+                                c("csv", "geojson"))
 
 # load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data <- filter(datastore_resources, row_number() == 1) %>% get_resource()
 data
 
 #### Save data ####
-write_csv(data, "data/raw_data/unedited_data.csv") 
+write_csv(data, "data/raw_data/unedited_data.csv")

### 2024-09-27 scripts/02-data_cleaning.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the city council attendance data 
+# Purpose: Cleans the city council attendance data
 # Author: Wendy Yuan
 # Date: 27 September 2024
 # Contact: w.yuan@mail.utoronto.ca
@@ -16,15 +16,14 @@ raw_data <- read_csv("data/raw_data/unedited_data.csv")
 cleaned_data <-
   raw_data |>
   janitor::clean_names() |>
-  filter(committee == "City Council") |> 
+  filter(committee == "City Council") |>
   # changing present to be more readable
   mutate(present = ifelse(present == "Y", "Yes", "No")) |>
   mutate(name = paste(first_name, last_name, sep = " ")) |>
   # separate year from session_date
-  mutate(session_year = substr(session_date, 1, 4)) |> 
+  mutate(session_year = substr(session_date, 1, 4)) |>
   # selects columns of interest
   select(name, committee, session_year, session_type, present)
 
 #### Save data ####
 write_csv(cleaned_data, "data/cleaned_data/cleaned_data.csv")
-

### 2024-09-27 scripts/03-test_data.R

@@ -7,8 +7,8 @@
 library(readr)
 
 # creating inital variables
-begin_year = 2022
-end_year = 2026
+begin_year <- 2022
+end_year <- 2026
 
 #### Tests raw data ####
 raw_data <- read_csv("data/raw_data/unedited_data.csv")
@@ -22,14 +22,16 @@ cleaned_data <- read_csv("data/cleaned_data/cleaned_data.csv")
 #Test first last name deliminator
 all(sapply(cleaned_data$name, function(name) {
   split_name <- strsplit(name, " ")[[1]]  # Split by space
-  length(split_name) == 2                 # Check if it splits into exactly 2 parts
+  length(split_name) == 2 # Check if it splits into exactly 2 parts
 }))
 
 # Test for City Council Member
 all(cleaned_data$committee == "City Council")
 
 # Test session year
-all(cleaned_data$session_year >= begin_year & cleaned_data$session_year <= end_year & !is.na(cleaned_data$session_year))
+all(cleaned_data$session_year >= begin_year &
+      cleaned_data$session_year <= end_year &
+      !is.na(cleaned_data$session_year))
 
 #Test for Session types
 all(cleaned_data$session_type %in% c("Morning", "Afternoon", "Evening"))
